---
title: "Homework-27"
author: "Anna"
date: "11/8/2022"
output: html_document
---

# INTRODUCTION
In this homework, I will be reading through chapters 1 and 2 ("Introduction" and "First Steps", respectively), of the ggplot2 guide. This guide discusses how to make "elegant graphics for data analysis".

# CHAPTER 1

## 1.1 Welcome to ggplot2
ggplot2 uses the "Grammar of Graphics" which means it combines individual components to create figures. This allows for greater customization! 

This system works with layers, starting with the raw data and then adding annotations and stat summaries. Compared to base R, ggplot2 more easily allows you to combine graphics with other data sets and plots. 

## 1.2 What is the grammar of graphics?
"A graphic maps the data to the aesthetic attributes (colour, shape, size) of geometric objects (points, lines, bars)" ... the plot can also include stats transformations, info about the plots' coordinate system, and facetting.

Five mapping components:

1. Layer - Geometric elements (geom) and stats transformations (stat). 

2. Scale - Map data values to aesthetic values (color, shape, size, legend, axes).

3. Coord - Coordinate system to place data on a graphic plane (includes axes and gridlines).

4. Facet - Breaks up data into subsets.

5. Theme - Controls the general display, like font size and background color.

**The grammar for ggplot2 does NOT suggest which graphic to use or describe interactive graphics. 

## 1.3 How does ggplot2 fit in with other R graphics?
- Base R only lets you draw on top of a plot, not delete or modify existing content. They are simple and fast, but have a limited scope.

- Grid graphics is more complex than Base R, but does not have any tools for creating statistical graphics

- Lattice graphics lacks a formal model 

... with this in mind, ggplot2 aims to combine the good things about lattice and base graphics to create a strong model for creating graphics. It uses a grid to draw its graphics, which controls the appearance. With ggplot2, it is easy to introduce statistical graphics in adddition to plots. 

## 1.4 About this book
Chapter 2 - introduces geoms, aesthetic mappings, and facetting.

Chapters 3-9 - explains how to use the basic toolbox to fix common issues with data visualization.

Chapters 10-12 - controlling scales (axes, legends)

Chapter 13-14 - explanation of layering grammar 

Chapter 18 - theming system and saving plots


# CHAPTER 2

## 2.2 Fuel Economy data

Let's learn the basics of creting quick and easy plots with ggplot2...

Import the mpg data included in ggplot:
```{r}
library(ggplot2)
mpg
```

Variable descriptions:

1. 'cty' and 'hwy' record mpg for city and highway driving

2. 'displ' is engine displacement in liters

3. 'drv' is the drivetrain (f = front wheel, r = rear wheel, 4 = four wheel)

4. 'model' is the car model

5. 'class' is the type of car (two seater, SUV, compact...)

## 2.3 Key Components

Every ggplot2 plot has 3 components:

1. data

2. aesthetic mappings

3. geom layer(s)

Example:
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```
Layers are added with a '+' symbol

You could also be more efficient by not naming x and y, instead using this code:
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point()
```
ggplot2 recognizes that most plots have x and y axes, so these are the first two arguments in aes().

## 2.4 Color, size, shape and other aes() attributes

Let's look at the color argument:
```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()
```
This changes the colors of the points based on class, so it is easier to see any patterns based on grouping for this variable.

Compare the differences between these plots:
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue"))
ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue")
```
Assigning a color outside of the aes() function changes the actual color of the points to the colors provided in R. 

*Color and shape work best for categorical variables, size works well for continuous data

## 2.5 Faceting

Two types of faceting: grid and wrapped

facet_wrap(~) is a useful function. It splits the data into groups based on he name you input after '~'

For example...
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  facet_wrap(~class)
```

## 2.6 Plot geoms

Create different kinds of plots by selecting different geom layers, such as:

geom_smooth() - fits a smoother to the data

geom_boxplot()

geom_histogram()

geom_bar() - for categorical variables

geom_path() - draws a line between data points

*You can add a smoother to a scatteplot by adding it as another layer:
```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth()
# to turn off the assessment of uncertainty, insert 'se = FALSE' into the geom_smooth() function
```



