---
title: "Homework-25.Rmd"
author: "Anna"
date: "11/3/2022"
output: html_document
---

# INTRODUCTION
## The following script investigates the second half of GSWR ch. 5, linear regression and one-way ANOVA. We will be looking at general linear models, not to be confused with generalized linear models (GLM). 
* One of the most important functions in this section is lm() which fits linear models. 

A quick reminder about the workflow for this chapter:
1. Plot
2. Model
3. Check assumptions
4. Interpret
5. Plot again

# Simple linear regression

Setting things up;
```{r}
rm(list=ls())
library(here)
library(tidyverse)
```

Importing the data:
```{r}
plant_gr <- read.csv(here("Data", "plant.growth.rate.csv"))
glimpse(plant_gr)
```

Model the data:
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  theme_bw() +
  ylab("Plant growth rate (mm/week)")
```
Just by looking at this figure, I can see that there is a positive relationship between these two variables. By looking at the spread of the data points, we can make an estimate that the slope is about 15 (rise is ~30, run is ~2).

Note: df = (# of data points) - (# of parameters/coefficients)

Now that I have a figure and have examined the data, I can make a simple linear regression using the lm() function:
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, data = plant_gr)
```

Review of steps so far:
1. Create a linear model with lm()
2. Import the ggfortify package
3. Use the autoplot() function

```{r}
library(ggfortify)
autoplot(model_pgr, smooth.colour = NA)
# smooth.colour = NA removes the squiggly lines, replacing them with normal straight lines for analysis
```

What do these graphs mean?
1. Residuals vs. Fitted - Tells us if a line fits the data. The example above shows even distribution of the residuals... "hump" or "valley" shapes on this graph would be bad news.
2. Normal Q-Q - Shows if data are normally distributed. Ddata points should fall along the dotted line.
3. Scale-Location - Evaluates assumption of equal variance. No pattern should be prevalent between the points.
4. Residuals vs. Leverage - Detects outliers/influential data points.

## Introducing anova() and summary()

* anova() does NOT perform an ANOVA! (comparison of means). anova() DOES produce a sums-of-squares table (f-value, r-squared, adjusted r-squared).
* summary() produces a table of the estimates of the coefficients of the "model" line (intercept and slope).

```{r}
anova(model_pgr) 
summary(model_pgr)
```
From GSWR ch.5, we can make the following cnclusion based on these tests...
"Soil moisture had a positive effect on plant growth. Fore each unit increase in soil moisture, plant growth rate increased by 12.7 mm/week (slope = 12.7, t = 12.5, df = 48, p < 0.001)."

Next, we need to incorporate the statistical model into our original model by adding a regression line:
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab("Plant growth rate (mm/week)") +
  theme_bw()
  
```
* This may not work perfectly for more complicated models!

# Analysis of variance: the one-way ANOVA











