---
title: "Homework-23.Rmd"
author: "Anna"
date: "10/24/2022"
output: html_document
---

Setting things up
```{r}
rm(list=ls())
library(here)
library(dplyr)
library(ggplot2)
```

Import data
```{r}
DF <- read.csv(here("Data", "soil_data.csv"))
glimpse(DF)
```

Make a new column with 'Field.name' values as just the site name
```{r}
DF <- DF %>%
  mutate(Site = case_when(
    startsWith(Field.Name, "DEGRASSE") ~ "Degrasse", 
    startsWith(Field.Name, "DONNERVILLE") ~ "Donnerville", 
    startsWith(Field.Name, "KIP") ~ "Kip Tract", 
    startsWith(Field.Name, "PEAVINE") ~ "Peavine", 
    startsWith(Field.Name, "SOUTH") ~ "South Hammond"
    ))
```


Making a scatterplot of the SOM_percent data based on site
```{r}
ggplot(DF, aes(x = Site, y = SOM_percent)) +
  geom_point() +
  xlab("Sample Site") +
  ylab("Soil Organic Matter %") +
  theme_bw()
```

Making a boxplot with SOM_percent data based on site
```{r}
ggplot(DF, aes(x = Site, y = SOM_percent)) +
  geom_boxplot() +
  xlab("Sample Site") +
  ylab("Soil Organic Matter %") +
  theme_bw()
```

Making a histogram with SOM_percent data
```{r}
ggplot(DF, aes(x = SOM_percent)) +
  geom_histogram() +
  xlab("Soil Organic Matter %")
```

Making a scatterplot with Ca.ppm data based on site
```{r}
ggplot(DF, aes(x = Site, y = Ca.ppm)) +
  geom_point() +
  xlab("Sample Site") +
  ylab("Soil Calcium (ppm)") +
  theme_bw()
```

Making a boxplot with Ca.ppm data based on site
```{r}
ggplot(DF, aes(x = Site, y = Ca.ppm)) +
  geom_boxplot() +
  xlab("Sample Site") +
  ylab("Soil Calcium (ppm)") +
  theme_bw()
```

Making a histogram with Ca.ppm data
```{r}
ggplot(DF, aes(x = Ca.ppm)) +
  geom_histogram() +
  xlab("Soil Calcium (ppm)")
```

