---
title: "Homework-24.Rmd"
author: "Anna"
date: "10/28/2022"
output: html_document
---

# INTRODUCTION
Always start data analysis with a figure before diving into the stats testing. This way, you can see expected results first and ensure that your statistical model is reliable

Workflow for GSWR ch. 5:
1. Plot your data 
2. Build a model
3. Check assumptions
4. Interpret the model
5. Replot the data and model

# 5.2 Chi-squared contingency table analysis
- Attempt to understand your hypothesis both biologically and statistically!
- Chi-squared is an analysis of COUNT DATA (two or more categorical variables)

Testing the null hypothesis that there is no association between frequency of black and red ladybirds...

First, set things up
```{r}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(here)
```

Import the ladybirds data
```{r}
lady <- read.csv(here("Data", "ladybirds.csv"))
glimpse(lady)
```

Test if the frequency of the two color morphs differ between two habitat sites. Start by organizing the data and calculating the totals for each color
```{r}
totals <- lady %>%
  group_by(Habitat, colour) %>%
  summarise(total.number = sum(number))
```

Use a bar chart* to summarize the raw data. This gives us a general idea of the spread of the data.
```{r}
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
# stat = 'identity' tells R not to try calculating anything from the data
# position = 'dodge' tells R to use grouping to separate the bars into designated categories
  scale_fill_manual(values = c(black = "black", red = "red"))
# scale_fill_manual() allows you to change the colors of the bars 
```
*For count data, each bar is one point and there is no real zero or variation to examine, which is why a bar graph is appropriate here. 

Making the chi-sq test:
To make it do a two-way contingency test, we need to provide a matrix...
```{r}
lady.mat <- xtabs(number ~ Habitat + colour, data = lady)
```

Then using this matrix, we can perform the chi-sq test:
```{r}
chisq.test(lady.mat)
```
The resulting p-value is less than 0.05, so we can reject the null hypothesis. There is in fact an association between Habitat and colour morph.

We can view the mechanics of the chi-sq test by assigning the chi-sq test outputs to a variable and looking at them...
```{r}
lady.chi <- chisq.test(lady.mat)
names(lady.chi)
lady.chi$expected
```

# 5.3 Two-sample t-test

A few points...
1. PLot your data!
2. Check model assumptions
3. R makes these two steps easy

What is the two-sample t-test? It compares the means of two groups of numeric values, and it can be used for groups with small sample sizes











