---
title: "Homework-21.Rmd"
author: "Anna"
date: "10/21/2022"
output: html_document
---

In this homework, I will be learning how to make basic figures using ggplot2.

Setting things up:
```{r}
rm(list=ls())
library(ggplot2)
library(dplyr)
library(here)
```

Import data:
```{r}
DF <- read.csv(here("Data", "compensation.csv"), stringsAsFactors = TRUE)
glimpse(DF) #take a look at the data
```

Making a basic scatterplot with the ggplot() function... The first argument in this function is the dataframe. aes() controls the aesthetics, such as specifying which variable is on the x and y axes.'+' adds layers to the graph
```{r}
ggplot(DF, aes(x = Root, y = Fruit, shape = Grazing)) + #adding the 'color' argument displays points as different colors to match levels in groups. 'shape' can be used in place of color to differentiate the points by shape instead
  geom_point(size = 5) + #adds the data as points, which we can determine the size of
  xlab("Root Biomass") + #labels the x-axis
  ylab("Fruit Production") + #labels the y-axis
  theme_bw() #gets rid of the grey background
```

Box-and-whisker plots...
```{r}
ggplot(DF, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() + #boxplot function within ggplot2
  geom_point(size = 4, color = 'light grey', alpha = 0.5) + #adds data points as a layer on the boxplot. 'alpha' changes transparency of points
  xlab("Grazing treatment") +
  ylab("Fruit production") +
  theme_bw()
```

Histograms...
```{r}
ggplot(DF, aes(x = Fruit)) +
  geom_histogram(bins = 10) + #you can change number of bins or 'binwidth' to make the histogram prettier. ex: binwidth = 15 would produce a similar graph to bins = 10 in this case
  facet_wrap(~Grazing) #faceting groups data based on a specified variable
```

How to save graphs...
  1. Plots tab > Export OR
  2. ggsave() function
The ggsave function lets you name your saved file within the function and if you specify file type (ex: .png), it will automaticaly know to save it as a .png file
Example...
```{r}
ggsave("ThatCoolHistogramOfFruit.png")
```
?ggsave is a good place to go for more info


THIS LOOKS GOOD BUT YOU SHOULD PROVIDE MORE THOUGHTS ABOUT YOUR CODE OUTSIDE OF CODE BLOCKS. FOR EXAMPLE, THOUGH IT MIGHT SEEM OBVIOUS NOW, FUTURE ANNA MIGHT NOT RECALL WHAT THE BOOK SAYS ABOUT HOW TO CHANGE THE BINS ON A GGPLOT (OR WHY) OR WHAT THE ADVANTAGES ARE TO A SCATTERPLOT, FOR INSTANCE.