---
title: "Homework-26.Rmd"
author: "Anna"
date: "11/2/2022"
output: html_document
---

Setting things up:
```{r}
rm(list=ls())
library(here)
library(tidyverse) # Uploads dplyr and ggplot2
library(ggfortify)
```

Importing the data:
```{r}
soil <- read.csv(here("Data", "soil_reduced-1.csv"), stringsAsFactors = T)
```

Before we begin any analysis, we need to check our data for possible errors/outliers. Let's begin with that.

A. Forest names
```{r}
levels(soil$Forest)
# All good, there are 5 different forests listed as expected
```

B. Soil Organic Matter and Soil Calcium
```{r}
#Start with a histogram for SOM %
ggplot(soil, aes(x = SOM_percent)) +
  geom_histogram()
```

Now check Soil Calcium
```{r}
ggplot(soil, aes(x = Ca_ppm)) +
  geom_histogram()
```

Let's also run a summary on all the variables:
```{r}
summary(soil)
```


## We'd like to ask the following questions for these data:

### 1. Is there a relationship between Soil organic matter (SOM, %) and soil calcium (Ca_ppm)?

#### Plot relationship
Both variables are continuous...
```{r}
ggplot(soil, aes(x = SOM_percent, y = Ca_ppm, color = Forest)) +
  geom_point() +
  theme_bw() +
  xlab("Soil Organic Matter %") +
  ylab("Calcium (ppm)")
```
From looking at the data, it appears there may be a weak positive relationship between SOM and Ca.

Can I guess the slope or y-int?
Slope is about 1000/53 (18ish) and the y-int is about -5

What stats model fits these data?
Both response and predictor are continuous, so linear regression seems appropriate

So let's construct the model:
```{r}
SOM_Ca_mod <- lm(Ca_ppm ~ SOM_percent, data = soil)
```

Now I need to check model assumptions before interpreting.
```{r}
autoplot(SOM_Ca_mod)
```
Based on my observations of the autoplot() results, I am modestly comfortable that my data meet the lm() assumptions. 

Now look at model outcome:
```{r}
anova(SOM_Ca_mod)
```
The p-value is greater than 0.05, so we accept the null hypothesis that there is no relationship between SOM and Ca.

2. Does the abundance of worms differ between Donnerville and the Kip Tract?

3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?

4. Does the abundance of zinc (Zn_ppm) differ among forests?




