---
title: "Homework-25.Rmd"
author: "Anna"
date: "11/3/2022"
output: html_document
---

# INTRODUCTION
## The following script investigates the second half of GSWR ch. 5, linear regression and one-way ANOVA. We will be looking at general linear models, not to be confused with generalized linear models (GLM). 
* One of the most important functions in this section is lm() which fits linear models. 

A quick reminder about the workflow for this chapter:
1. Plot
2. Model
3. Check assumptions
4. Interpret
5. Plot again

# Simple linear regression

Setting things up;
```{r}
rm(list=ls())
library(here)
library(tidyverse)
```

Importing the data:
```{r}
plant_gr <- read.csv(here("Data", "plant.growth.rate.csv"))
glimpse(plant_gr)
```

Model the data:
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  theme_bw() +
  ylab("Plant growth rate (mm/week)")
```
Just by looking at this figure, I can see that there is a positive relationship between these two variables. By looking at the spread of the data points, we can make an estimate that the slope is about 15 (rise is ~30, run is ~2).

Note: df = (# of data points) - (# of parameters/coefficients)

Now that I have a figure and have examined the data, I can make a simple linear regression using the lm() function:
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, data = plant_gr)
```

Review of steps so far:
1. Create a linear model with lm()
2. Import the ggfortify package
3. Use the autoplot() function

```{r}
library(ggfortify)
autoplot(model_pgr, smooth.colour = NA)
# smooth.colour = NA removes the squiggly lines, replacing them with normal straight lines for analysis
```

What do these graphs mean?
1. Residuals vs. Fitted - Tells us if a line fits the data. The example above shows even distribution of the residuals... "hump" or "valley" shapes on this graph would be bad news.
2. Normal Q-Q - Shows if data are normally distributed. Ddata points should fall along the dotted line.
3. Scale-Location - Evaluates assumption of equal variance. No pattern should be prevalent between the points.
4. Residuals vs. Leverage - Detects outliers/influential data points.

## Introducing anova() and summary()

* anova() does NOT perform an ANOVA! (comparison of means). anova() DOES produce a sums-of-squares table (f-value, r-squared, adjusted r-squared).
* summary() produces a table of the estimates of the coefficients of the "model" line (intercept and slope).

```{r}
anova(model_pgr) 
summary(model_pgr)
```
From GSWR ch.5, we can make the following cnclusion based on these tests...
"Soil moisture had a positive effect on plant growth. Fore each unit increase in soil moisture, plant growth rate increased by 12.7 mm/week (slope = 12.7, t = 12.5, df = 48, p < 0.001)."

Next, we need to incorporate the statistical model into our original model by adding a regression line:
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab("Plant growth rate (mm/week)") +
  theme_bw()
  
```
* This may not work perfectly for more complicated models!

# Analysis of variance: the one-way ANOVA

For this kind of test, the explanatory variable is categorical. In the following example, we will analyze a data frame with multiple levels of a variable.

Setting things up:
```{r}
rm(list=ls())
library(here)
library(ggplot2)
library(dplyr)
library(ggfortify)
```

Import the data:
```{r}
daphnia <- read.csv(here("Data", "Daphniagrowth.csv"))
glimpse(daphnia)
# glimpsing the data shows me that I have character, integer, and decimal variables
```

First, model the data with a boxplot to view the variation in the response variable as a function of grouping.
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() # this switches the x and y axes
```
Just by looking at this model, there appears to be a difference in growth rate based on different types of parasites. The control treatment appears to produce the highest growth rate.

Time to construct the linear model... We will use lm() to perform the one-way ANOVA
```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

Just like with the linear regression model, we make 4 diagnostic graphs using autoplot()
```{r}
autoplot(model_grow, smooth.colour = NA)
```

We can then draw conclusions using anova() and summary()
```{r}
anova(model_grow)
```
This shows us that different parasite treatments have produced an effect on growth rate. The small p-value allows us to reject the null hypothesis that there are no differences between treatment groups. The F-value is a ratio (between-group variance : within-group variance) so a large F-value produces a small p-value.

Treatment contrasts... show the difference between the reference level (control) and the other treatmenys
```{r}
summary(model_grow)
```
*R defaults to representing things in alphabetical order. 




