---
title: "Applying more dplyr to my data"
author: "Anna"
date: "9/28/2022"
output: html_document
---

My goal for this homework is to begin to feel more comfortable using the dplyr functions from GSWR chapters 2 & 3. This includes subsetting, manipulating, and renaming data, as well as making new data frames.

Setting things up
```{r}
rm(list=ls())
library(here)
library(dplyr)
library(ggplot2)
```

Importing data from a .csv file and viewing it in a new tab
```{r}
DF <- read.csv(here("Data", "day_one_data_for_homework13.csv"))
View(DF)
```

Summarizing the data
```{r}
summary(DF)
```

Renaming column headers
```{r}
names(DF) #copy the names from this output
DF <- rename(DF, "Tree_num" = Tree..,"Species" = Tree.species, "leaf_length" = Leaf.length,"leaf_width" = leaf.width)
```


PROBLEM #1: leaf width and length unit discrepancy
Fixing leaf measurements (first, grab data of interest)
```{r}
to_fix <- DF %>%
  filter(leaf_units == "inches") %>%
  select(c(leaf_length, leaf_width))
View(to_fix)
```
Then, add two variables to this data frame by converting each measurement to a decimal (by hand or on paper)
```{r}
to_fix <- to_fix %>%
  mutate(leaf_length_numeric = c(4.125, 6.938, 4.5, 3.75, 6))
to_fix <- to_fix %>%
  mutate(leaf_width_numeric = c(3.938, 6.5, 6.125, 3.0, 2.5))
```
Now, convert to cm
```{r}
to_fix <- to_fix %>%
  mutate(leaf_length_cm = leaf_length_numeric*2.54, leaf_width_cm = leaf_width_numeric*2.54)
```
Now, we can put those values into DF using indexing
```{r}
DF$leaf_length[1:5] <- to_fix$leaf_length_cm[1:5]
DF$leaf_width[1:5] <- to_fix$leaf_width_cm[1:5]
```
Change units for these new rows so they are cm
```{r}
DF$leaf_units[1:5] <- "cm"
```
Last step is to make leaf length and width as numeric variables
```{r}
DF$leaf_length <- as.numeric(DF$leaf_length)
DF$leaf_width <- as.numeric(DF$leaf_width)
```

PROBLEM #2: Species
```{r}
unique(DF$Species) #all the species are listed in different ways and therefore are considered all unique strings
```
Find all the rows with maple in the name. Use the which() function to return row numbers
```{r}
maples <- which(DF$Species == "red maple" | DF$Species == "norway maple" | DF$Species == "Maple" | DF$Species == "maple" | DF$Species == "Red maple")

DF$Species[maples] <- "maple species"
```



New data frame with a subset from the original DF
```{r}
English <- DF %>%
  filter(leaf_units == "inches")
```

New data frame with dates correctly formatted. Use the filter() function to select certain observations.
```{r}
Uniform_dates <- DF %>%
  filter(Date == "8/24/22")
```

