---
title: "Wolke_HW11_GSWRCh3Ex1.Rmd"
author: "Anna"
date: "9/25/2022"
output: html_document
---

Chapter 3 in GSWR is dedicated to data manipulation within R using the `dplyr` package. Let's dive into the fundamentals of data visualization and analysis...

Setting things up
```{r}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(here)
```

Import the data set
```{r}
DF <- read.csv(here("Data", "compensation.csv"))
glimpse(DF) #to check the structure of the data, learned in the last chapter
```

#3.1 SUMMARY STATS FOR EACH VARIABLE
Using the `summary()` function allows you to see a summary of the stats of a data set, including
minimum, IQR, median, mean, and maximum for the numeric columns. Levels and sample size are given for categorical columns
```{r}
summary(DF)
```

#3.2 DPLYR VERBS
More helpful functions found in the `dplyr` package
```{r}
select() #selecting columns
slice() #selecting rows
filter() #selecting subsets of rows based on column values
arrange() #sorting rows
mutate() #creating new variables in a data frame
```
Key to remember: The first argument to all dplyr functions is the data frame (gives the function a reference to which data set you want to manipulate)

#3.3 SUBSETTING
Example of how to select the Fruit column in the compensation data
```{r}
select(DF, Fruit)
```
This asks for a part of a data frame, and returns a data frame to you
Note: All dplyr functions only do one thing very quickly

Example of how to select all columns EXCEPT the Root column
```{r}
select(DF, -Root) #add a minus sign to leave out a column from the returned data frame
```

Example of how to grab certain rows
```{r}
slice(DF, 2) #grabs a single row
slice(DF, 2:10) #the : allows selection of a series of rows
slice(DF, c(2,3,10)) #discontinuous sets are "collected" with the c() function
```
Note: the returned data frame does not display the row number identity from the original data frame

Logical operators and booleans:
```{r}
with(DF, Fruit > 80)
filter(DF, Fruit > 80)
filter(DF, Fruit > 80 | Fruit < 20)
```
The logical operators/booleans are simple things you can ask R so it returns true/false data. Some logical operators include...
`==` equals
`!=` does not equal
`<` less than
`>` greater than
`>=` greater than or equal to
`<=` less than or equal to
`|` or (between two logical operators)
`&` and (between two logical operators)

To use the outputs of these outputs in subsequent calculations/analysis, assign them to a variable
```{r}
lo_hi_fruit <- filter(DF, Fruit > 80 | Fruit < 20)
lo_hi_fruit #run the variable to see the output
```

#3.4 TRANSFORMING
Example of using the mutate() function to create a new column in the data frame:
```{r}
DF <- mutate(DF, logFruit = log(Fruit))
head(DF)
```
Note: By working with data in R, we are never manipulating the original raw data because it is saved as a separate .csv file on our computer

#3.5 SORTING 
The arrange() function allows us to arrange rows in an order that makes sense, such as increasing Fruit production for the compensation data
```{r}
arrange(DF, Fruit)
```

#3.6 MINI-SUMMARY AND TIPS
Top Tip 1: You can use more than one dplyr function in one line of code
```{r}
select(filter(DF, Fruit > 80), Root)
```

Top Tip 2: Use pipe commands! %>% means "put the answer of the left-hand command into the function on the right" and always start with the data frame
```{r}
DF %>%
  filter(Fruit > 80) %>%
    select(Root)
#Shows the values of the Root column associated with the values of the Fruit columns where the fruit production is greater than 80
```

#CALCULATING SUMMARY STATS
Method 1: Nesting
```{r}
mean.fruit <- summarize(
  group_by(DF, Grazing), 
    meanFruit = mean(Fruit))
```

Method 2: Piping
```{r}
DF %>%
  group_by(Grazing) %>%
    summarize(
      meanFruit = mean(Fruit),
      sdFruit = sd(Fruit))
    
```



