---
title: "Wolke-Homework19.Rmd"
author: "Anna"
date: "10/16/2022"
output: html_document
---

# Introduction
For this homework assignment, my goal is to get more comfortable with the basics of RStudio (edittin)

Setting things up:
```{r}
rm(list=ls()) #clears R's brain
library(dplyr) #imports dplyr package
library(here) #allows use of the here() function while importing data from a file
```

Import the data using the here() function and name it "Leaves":
```{r}
Leaves <- read.csv(here("Data", "Class_leaf_measurements.csv"), stringsAsFactors = TRUE) #reads the .csv file and makes it into a data frame
```

Examine the new dataset with the str() function
```{r}
str(Leaves) #shows the structure of the data frame
```
This specifies the factor data because I used stringsAsFactors = TRUE while importing the data

Renaming column headers:
```{r}
Leaves <- rename(Leaves, "leaf_num" = Leaf.., "DBH_cm" = DBH..cm., "petiole_length" = Petiole.length) #for the rename() function, first list the DF, then put the new column header names in parentheses, equal to the old column header name
```

Since I already included stringsAsFactors = TRUE while importing the data,. Species is already a factor, and using the str() function shows that it has 9 levels. There should only be 3 species, so this tells me that there is inconsistancy in how the species names were entered into the original data. Time to fix that...
```{r}
unique(Leaves$Species) # this shows the unique levels within the Species data
norway <- which(Leaves$Species == "Norway maple" | Leaves$Species == "Norway Maple" | Leaves$Species == "Norway mapel")
sugar <- which(Leaves$Species == "sugar maple" | Leaves$Species == "Sugar Maple" | Leaves$Species == "Sugar maple")
red <- which(Leaves$Species == "red maple" | Leaves$Species == "Red Maple" | Leaves$Species == "Red maple")
```
And then to make the official fix...
```{r}
Leaves$Species[norway] <- "Norway maple"
Leaves$Species[sugar] <- "sugar maple"
Leaves$Species[red] <- "red maple"
```

Making a new data frame "Sugar" to isolate the sugar maple data from the rest:
```{r}
Sugar <- Leaves %>%
  select(-Person, -Tree_num, -leaf_num) %>%# to take out certain columns, use the select() function
  filter(Species == "sugar maple") # this filters out all species that are not sugar maple
```


Now, we want to create a histogram of petiole lengths based on the Sugar data frame:
```{r}
hist(Sugar$petiole_length, main = "Histogram for Petiole Lengths", xlab = "Petiole length")
# xlab changes the label on the x-axis
# main changes the title
```
I'M GLAD YOU FIGURED OUT HOW TO PLOT A HISTOGRAM. CHAPTER 2 OF THE BOOK HAS YOU USE GGPLOT. HERE'S SOME CODE TO TRY A "TIDYVERSE" APPROACH TO THE PLOT. IT REQUIRES GGPLOT, WHICH YOU DID NOT TURN ON ABOVE, BUT SHOULD HAVE

```{r}
library(ggplot2)
ggplot(Sugar, aes(x = petiole_length))+
  geom_histogram()
```


